---
title: "Getting Started with {leaf.magic}"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{leaf-magic}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%",
  message = FALSE,
  warning = FALSE
)
```

```{r setup, warning=FALSE}
library(leaflet)
library(leaf.magic)
```

# Purpose

The purpose of `{leaf.magic}` is to extend R `{leaflet}`'s capability to use modern icon sets. While `leaflet::addAwesomeMarkers()` exists, it has a few core disadvantages; markers are a fixed size, can only be one of 19 colours, and pull icons from significantly out of date icon sets.

`{leaf.magic}` uses contemporary R packages - currently `{fontawesome}` and `{bsicons}` - to access up-to-date SVG icon sets, `{magick}` to knit them onto map markers, and then `leaflet::makeIcon()` to translate them into a `{leaflet}`-usable map marker.

Some core advantages of `magicIcons()` over `leaflet::awesomeIcons()` are:

* Markers pull from up-to-date, larger icon sets.

* Markers and icons can be any colour, and can be resized. 

One *disadvantage* is that `magicIcons()` may be slower initially to create large numbers of different icons. However, once a marker has been created in an R session, it will be significantly quicker to redraw, closing the gap between `awesomeIcons()` and `magicIcons()`.

# Example Data

In this document we'll use the in-built `port_talbot` dataset, which details the location of some air quality measurement stations around the town of Port Talbot, Wales, UK, some of which are open and some of which have closed. 

We can give this data a look using vanilla `{leaflet}`:

```{r map-basic}
port_talbot$open_year <- as.integer(format(port_talbot$start_date, "%Y"))

port_talbot$popup <-
  paste0(
    "<strong>",
    toupper(port_talbot$site),
    "</strong> (",
    port_talbot$code,
    ")<hr>Site Type: ",
    port_talbot$site_type,
    "<br>Opened: ",
    port_talbot$open_year
  )

leaflet(port_talbot) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addMarkers(popup = ~popup)
```

# Simple Example

We can use the `magicIcons()` function to swap out the default markers for some `{leaf.magic}` markers. The important argument here is `icon`, which is the (in this case) Font Awesome icon of interest^[See `fontawesome::fa_metadata()` for a complete list].

```{r map-simple}
leaflet(port_talbot) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addMarkers(
    popup = ~popup,
    icon = magicIcons(
      icon = "cloud"
    )
  )
```

# Varying Icons

`{leaf.magic}` allows you to vary icons by some other variable. You can do this manually, or use one of `{leaf.magic}` constructor functions like `iconFactor()`.

```{r map-varicon}
site_types <- c("Urban Industrial", "Urban Background", "Urban Traffic")

iconPal <- iconFactor(
  icons = c("industry", "house", "car"),
  levels = site_types
)

leaflet(port_talbot) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addMarkers(
    popup = ~popup,
    icon = ~ magicIcons(
      icon = iconPal(site_type)
    )
  )
```

Note that there's also `iconBin()` and `iconQuantile()` which can help map icons to numeric data. In this case, it makes most sense to choose icons that have some inherent "order" to them.

These functions don't have all of the features of `colorBin()` and `colorQuantile()`, in part due to the differences between mapping a continuous aesthetic like colour and a discrete aesthetic like icons. A nice feature is that these are very much icon-led - the `breaks` and `n` arguments default to the length of the provided `icons`, so you don't need to specify them if you're happy letting the functions decide the break-points/`probs` for you.

```{r map-varicon-q}
iconPalYr <- iconQuantile(
  icons = c("hourglass-start", "hourglass-half", "hourglass-end"),
  domain = port_talbot$open_year
)

leaflet(port_talbot) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addMarkers(
    popup = ~popup,
    icon = ~ magicIcons(
      icon = iconPalYr(open_year)
    )
  )
```

# Varying Colours

`magicIcons()` has two arguments related to colour; `markerColor` which colours the tear-drop-shaped marker, and `iconColor` which colours the icon. Any hex-code can be supplied to these arguments - either a constant value or colour values mapped to a column using `leaflet::colorFactor()`.

```{r map-varcol-fct}
catPal <- colorFactor(c("royalblue", "tomato", "gold"), levels = site_types)

leaflet(port_talbot) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addMarkers(
    popup = ~popup,
    icon = ~ magicIcons(
      icon = iconPal(site_type),
      iconColor = catPal(site_type),
      markerColor = "white"
    )
  ) %>%
  addLegend(
    pal = catPal,
    values = port_talbot$site_type,
    title = "Site Type"
  )
```

As the markers themselves can be set to *any* colour, we can even use a continuous scale (e.g., from `leaflet::colorNumeric()`).

```{r map-varcol-num}
palNum <- colorNumeric("viridis", domain = port_talbot$open_year)

leaflet(port_talbot) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addMarkers(
    popup = ~popup,
    icon = ~ magicIcons(
      icon = iconPalYr(open_year),
      markerColor = palNum(open_year),
      iconColor = "white"
    )
  ) %>%
  addLegend(
    pal = palNum,
    values = port_talbot$open_year,
    title = "Opening Year",
    labFormat = labelFormat(big.mark = "")
  )
```

# Varying Size

Unlike `leaflet::awesomeIcons()`, `magicIcons()` can be easily re-sized. Much like in `leaflet::addCircleMarkers()`, this can vary with another variable; let's make the closed sites a bit smaller than the open ones with `markerSize`. Note the default is `30L`, which is roughly the same size as the default `leaflet::addMarkers()` marker.

```{r map-varsize}
leaflet(port_talbot) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addMarkers(
    popup = ~popup,
    icon = ~ magicIcons(
      icon = iconPal(site_type),
      markerColor = "white",
      markerSize = ifelse(open, 30L, 20L),
      iconColor = catPal(site_type)
    )
  ) %>%
  addLegend(
    pal = catPal,
    values = port_talbot$site_type,
    title = "Site Type"
  )
```


# Icon Legends

`{leaf.magic}` provides the `addIconLegend()` function to help construct an icon legend, similar to `leaflet::addLegend()` constructs colour legends.

```{r map-iconlegend}
leaflet(port_talbot) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addMarkers(
    popup = ~popup,
    icon = ~ magicIcons(
      icon = iconPal(site_type),
      markerColor = palNum(open_year),
      iconColor = "white"
    )
  ) %>%
  addLegend(
    pal = palNum,
    values = port_talbot$open_year,
    title = "Opening Year",
    labFormat = labelFormat(big.mark = "")
  ) %>%
  addIconLegend(
    icons = c("industry", "car", "house"),
    labels = site_types,
    title = "Site Type"
  )
```

Often, the colour of the marker (or icon) and the icon itself will align. In that case, the `colors` argument of `addIconLegend()` can be used to create an efficient, combined legend.

```{r map-iconcollegend}
leaflet(port_talbot) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addMarkers(
    popup = ~popup,
    icon = ~ magicIcons(
      icon = iconPal(site_type),
      markerColor = "white",
      iconColor = catPal(site_type)
    )
  ) %>%
  addIconLegend(
    icons = iconPal(site_types),
    labels = site_types,
    colors = catPal(site_types),
    title = "Site Type"
  )
```

# Other Utilities

`{leaf.magic}` exports the `awesomePalette` list, which are hex codes for the colours used in `leaflet::awesomeIcons()`. You could use these if you want to recreate the colour scheme of `awesomeIcons()` with the flexibility of `magicIcons()`.

```{r awesomepal}
cols <- names(awesomePalette)[names(awesomePalette) != "white"]

breweries91$facolor <- sample(cols, nrow(breweries91), replace = TRUE)

breweries91$hexcolor <- unlist(use.names = FALSE, awesomePalette[breweries91$facolor])

leaflet(breweries91) %>%
  addTiles() %>%
  addAwesomeMarkers(
    icon = ~ awesomeIcons(
      icon = "circle",
      markerColor = facolor,
      iconColor = "#FFFFFF",
      library = "fa"
    ),
    group = "Awesome"
  ) %>%
  addMarkers(
    icon = ~ magicIcons(
      icon = "fas fa-circle",
      markerColor = hexcolor,
      iconColor = "#FFFFFF"
    ),
    group = "Magic"
  ) %>%
  addLayersControl(
    baseGroups = c("Awesome", "Magic")
  )
```

